---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# inlineOutputAddin

<!-- badges: start -->
<!-- Add badges here later (CRAN, R-CMD-check, etc.) -->
<!-- badges: end -->

`inlineOutputAddin` provides an RStudio addin that evaluates selected R code and inserts or updates a marked, commented output block directly in an `.R` script.

This makes plain R scripts behave a bit like lightweight notebooks, while remaining standard `.R` files.

## Installation

You can install the development version from GitHub:

```{r eval=FALSE}
# install.packages("pak")
pak::pak("YOUR_GITHUB_USERNAME/inlineOutputAddin")
```

Or using `remotes`:

```{r eval=FALSE}
remotes::install_github("YOUR_GITHUB_USERNAME/inlineOutputAddin")
```

## What It Does

Given code like:

```r
head(mtcars)
```

Running the addin inserts (or updates) a marked output block:

```r
head(mtcars)
# >>> output
#                    mpg cyl disp  hp drat    wt  qsec vs am gear carb
# Mazda RX4         21.0   6  160 110 3.90 2.620 16.46  0  1    4    4
# ...
# <<< output
```

If the code changes and the addin is run again, the existing output block is automatically replaced.

## How to Use

1. Open an `.R` script in RStudio.
2. Select code (or place the cursor on a single line).
3. Run the addin:

   - Via **Addins → Update Inline Output**
   - Or assign a keyboard shortcut in:
     `Tools → Modify Keyboard Shortcuts`

The function evaluates the code in the global environment and:

- Inserts a new output block if none exists
- Replaces the existing `# >>> output` / `# <<< output` block if present

## Output Markers

The addin uses the following markers:

```
# >>> output
# <<< output
```

Everything between them is treated as generated output and may be replaced on the next run.

## Notes

- The addin requires **RStudio**.
- It captures printed console output.
- Long outputs are truncated for readability.
- It does not modify code outside the marked output block.

## Example (Non-interactive)

Because this functionality depends on RStudio, examples are guarded:

```{r}
if (rstudioapi::isAvailable()) {
  # dscript()
}
```

## Motivation

Plain `.R` scripts are often preferred in institutional or production settings.  
This addin provides a lightweight way to keep code and results together without switching to R Markdown or Quarto.

## License

MIT © Daniel Rakotomalala
